---
title: "Palmer Penguins"
output: github_document
---
## TidyTuesday [2020-07-28]
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE, 
  fig.width = 10,
  fig.out="100%"
  )
```

```{r, warning=FALSE, message=FALSE}
library(palmerpenguins)
library(tidyverse)
```

A quick look at the tidied data set
```{r}
palmerpenguins::penguins
```

## Measurements split by sex and species
```{r}
palmerpenguins::penguins %>%  
  gather(k, v, -species, -island, -sex, -year) %>% 
  ggplot(aes(v))+
  geom_density(aes(y=..scaled.., fill=sex), alpha=1/2)+
  facet_grid(species ~ k, scales="free")+
  scale_fill_viridis_d()
```

* Males bigger than females
* Some observations with NA sex


## Measurements split by island and species
```{r}
palmerpenguins::penguins %>%  
  gather(k, v, -species, -island, -sex, -year) %>% 
  ggplot(aes(v))+
  geom_density(aes(y=..scaled.., fill=species), alpha=1/2)+
  facet_grid(island ~ k, scales="free")+
  scale_fill_viridis_d()
```

* Adelie on all three islands and no difference in distributions between islands. 
* Chinstrap unique to Dream Island and Gentoo unique to Biscoe. 
* Gentoo bigger than Adelie (and biggest overall) **BUT** smaller bill depth!
* Chinstrap similar to Adelie but much bigger bill length!

## Classifier
Build classifier from **all data** just for fun (I am not training a model and evaluating performance here!)
```{r}
rpart::rpart(data = palmerpenguins::penguins %>% select(-island),
             formula = species ~ .,
             model = TRUE) %>% 
  rpart.plot::rpart.plot(type=5)
```

* These splits appear sensible based on a visual inspection of the density distributions above
